---
title: "Data 200: Lab 3 - Part 2 Solutions"
subtitle: "Data visualization foundations"
author: "Catalina Medina"
format: 
  html:
    echo: false
    message: false
    warning: false
    toc: true
    embed-resources: true
---

In this lab we will use the `movies` dataset from the `openintro` package.

```{r}
library(openintro)
library(tidyverse)

data("movies")
```

## Question 1

Look at the help documentation for the `movies` data. What year were these movies released in?

**Solution:**

The movies are all from 2003.


## Question 2

Use `glimpse()` on the `movies` data. How many movies are in the dataset?

**Solution:**


## Question 3

Fix the code below to produce a box plot for box office earnings. Once you fix the code you can remove the `#| eval: false` from the code chunk (it prevents the code chunk from being run when the document is rendered).

```{r}
#| eval: false
ggplot(movies, aes(x = )) +
  geom_boxplot()
```


**Solution:**

```{r}
ggplot(movies, aes(x = box_office)) +
  geom_boxplot()
```

## Question 4

Produce the plot from part 1 of the lab to investigate movie ratings by genre.

**Solution:**

Both of these variables are categorical so we want a dodged bar plot (referring to question 8 from lab 3 part 1).

```{r}
ggplot(movies, aes(x = genre, fill = rating)) +
  geom_bar(position = "dodge")
```

Notice how the plot is pretty hard to read? I can choose to move genre to the y-axis to make it easier to read.

```{r}
ggplot(movies, aes(y = genre, fill = rating)) +
  geom_bar(position = "dodge")
```

## Question 5

Produce a plot to visualize the relationship between critic score and box office earning.

**Solution:**

I did not explicitly tell you what plot to make here, so I want you to figure this out.
First, identify what type of variable critic score is and what type box office earnings is.
Second, identify what type of plot would be appropriate for those variables.
Then you can write the code to produce the plot.

These are both numeric variables so a scatterplot could be appropriate.
```{r}
ggplot(movies, aes(x = score, y = box_office)) +
  geom_point()
```


## Question 6

Do you see any relationship between critic score and box office earning? Another way to think about this is "If someone told me a movie's critic score, would that give me some information about what the box office earnings could be?"

**Solution:**

Yes, there appears to be a weak relationship between them. As one increases the other seems to generally increase.

## Question 7

Pose your own question about this data. Produce an appropriate plot to help answer your question.

**Solution:**

I chose to ask: What is the distribution of critic score overall?

Critic score is a numerical variable, so I could use a histogram or box plot to look at this.

```{r}
ggplot(movies, aes(x = score)) +
  geom_histogram()
```


## Version control

Once you complete the lab make sure to save your file, commit your changes with an informative commit message, and push your changes to GitHub.
