---
title: ""
format: 
  pdf:
    echo: false
    message: false
    warning: false
    geometry:
      - bottom=1in
header-includes:
  - \usepackage{lastpage}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyhf{}
  - \fancyhead[L]{Data 200 Spring 2026 -- \textbf{Lab 15a}}
  - \fancyhead[R]{Normal distributions}
  - \fancyfoot[C]{\thepage\ of \pageref{LastPage}}
  - \renewcommand{\headrulewidth}{0.4pt}
---

```{r}
#| echo: false
library(patchwork)
library(knitr)
library(tidyverse)
library(openintro)
```

```{r}
plot_normal_prob <- function(value, mean, sd, end = 3, direction = "lower"){
  
  x <- seq(mean - (end * sd) , mean + (end * sd), 0.01)
  
  y <- dnorm(x, mean = mean, sd = sd)
  
  data <- data.frame(x = x, y = y)
  
  p <- ggplot(data = data, aes(x)) +
    stat_function(
      fun = stats::dnorm,
      n = 101,
      args = list(mean = mean, sd= sd)
    ) 
  
  if(direction == "lower"){
    
    p <- p +
      stat_function(
        fun = dnorm, 
        args = list(mean, sd), 
        geom = "area", 
        fill = "#D55E00", 
        alpha = 0.8,
        xlim = c(mean - (end*sd), value)
      ) +
      labs(
        caption = paste0(
          "The area under the curve is ",
          round(pnorm(value, mean, sd), digits = 4) * 100, 
          "%."
        )
      )
  }
  
  
  if(direction == "upper"){
    
    p <- p +
      stat_function(
        fun = dnorm, 
        args = list(mean, sd), 
        geom = "area", 
        fill = "#D55E00", 
        alpha = 0.8, 
        xlim = c(value, mean+(end * sd))
      ) +
      labs(
        caption = paste0(
          "The area under the curve is ", 
          round(pnorm(value, mean, sd, lower.tail = FALSE), digits = 4) * 100, 
          "%."
        )
      )
  }
  
  p
  
}

```


\newcommand{\blank}{\rule{2.5cm}{0.15mm}}

**Names**:

**Question 1**\newline
Below are the normal distributions of four variables. 
```{r}
set.seed(123)

library(tidyverse)

# Define your contexts with realistic parameters
curves <- tibble(
  mean = c(80, 67, 0, 16),  # Ragdoll kitten birthweight (ounces)
  sd = c(4, 5, 1.5, 2),     # kitten birthweight variability (oz)
  variable = c(
    "Temp",
    "Height",
    "Measurement error",
    "Ragdoll"
  )
) |> 
  mutate(variable = factor(
    variable,
    levels = c("Ragdoll", "Temp", "Measurement error", "Height"),
    labels = c("A", "B", "C", "D")
  )) |> 
  arrange(variable)

# Generate curve data
curve_data <- curves %>%
  group_by(variable, mean, sd) %>%
  reframe(
    x = seq(mean - 4 * sd, mean + 4 * sd, length.out = 300),
    y = dnorm(x, mean, sd)
  )

# Plot (students only see A–D)
ggplot(curve_data, aes(x, y)) +
  geom_line(linewidth = 1.2) +
  facet_wrap(~ variable, scales = "free") +
  labs(
    x = "Value",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)
```

**Question 1a**\newline
Identify the distribution plot for each variable.

- Camarillo's daily high temperature in July (in Fahrenheit)
- Height of students at CSUCI (in inches)
- Error in measuring the same object repeatedly (in mm)
- Birth weight of ragdoll kittens (in ounces)

**Question 1b**\newline
Match each variable/curve pair to the following forms of the normal distribution.

- Normal($\mu$ = 0, $\sigma$ = 1.5)
- Normal($\mu$ = 80, $\sigma$ = 4)
- Normal($\mu$ = 67, $\sigma$ = 5)
- Normal($\mu$ = 16, $\sigma$ = 2)

**Question 1c**

\vspace*{-0.5cm}

| Variable | Plot | Distribution | Mean | Std. Dev. |
|:----------------|:---|:-----|:---|:---|
| Camarillo's daily high temp. in July (in F) | | Normal | | |
| Height of students at CSUCI (in inches) | | Normal | | |
| Measurement error | | Normal | | |
| Birth weight of ragdoll kittens (in ounces) | | Normal | | |

: Fill in the table for each variable. {.striped .hover}


**Question 1d**

Approximate each probability

- Probability of a positive measurement error:

- Probability of CSUCI student being over 8 feet tall:

- Probability of a ragdoll kitten weighing less than 20 oz at birth:

- Probability of a daily high temperature for Camarillo in July being between 70 and 90 degrees Fahrenheit:

## Question 2

Which of the following variables do you think could be appropriately approximated by the normal distribution?

```{r}
#| echo: false
#| fig-width: 7
#| fig-height: 5
# -----------------------------------------------
# Part 2: Histograms: Which look approximately normal?
# -----------------------------------------------

n <- 500

hist_data <- tibble(
  salary = rlnorm(n, meanlog = log(65000), sdlog = 0.4),
  wait_time = runif(n, min = 0, max = 60),
  body_temp = rnorm(n, mean = 98, sd = 1),
  commute_time = c(rnorm(n / 2, 20, 5), rnorm(n / 2, 50, 7))
) %>%
  pivot_longer(
    everything(),
    names_to = "variable",
    values_to = "value"
  )

ggplot(hist_data, aes(value)) +
  geom_histogram(bins = 30, color = "white") +
  facet_wrap(~ variable, scales = "free", ncol = 2) +
  labs(
    x = "Variable",
    y = "Count"
  ) +
  theme_minimal(base_size = 14)
```

