---
title: "Data 200: Lecture 15b"
subtitle: "Poisson Distribution"
author: "Medina - CSU Channel Islands"
format: 
  revealjs:
    theme: [sky, ../slide-style.scss]
    slide-number: true
execute:
  echo: true 
---


```{r}
#| echo: false
options(scipen = 999)
library(janitor)
library(patchwork)
library(tidyverse)
theme_set(theme_bw(base_size = 22))
```


# Review: Normal distribution

## Normal distribution: Shape

```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 2
#| fig-align: center

data.frame(x = seq(-4, 4, length = 100)) |> 
  ggplot(aes(x = x)) +
  geom_function(fun = dnorm, args = list(mean = 0, sd = 1), color = "blue") +
  theme(
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()
  ) +
  labs(y = "Density", x = "Values")
```

Model for continuous numerical variables with:

- "bell shaped" distribution (symmetric with single peak)
- range is unrestricted, but low probability for values far from mean

## Normal distribution: Parameters

$$X \sim Normal(\mu, \sigma)$$

^ This mean "X follows a normal distribution with mean $\mu$ and standard deviation $\sigma$"

- $\mu$ = mean
- $\sigma$ = standard deviation

## Normal distribution: Examples

Recall in lab we explored potential normal distributions for:

- Measurement error $\sim N(\mu = 0, \sigma = 1.5)$
- Birth weight of ragdoll kittens (oz) $\sim N(\mu = 16, \sigma = 2)$
- Height of students at CSUCI (inches) $\sim N(\mu = 67, \sigma = 5)$
- Camarillo's daily high temp in July (°F) $\sim N(\mu = 0, \sigma = 1.5)$

# Poisson distribution

## Posson distribution

Model for counts over a fixed interval (positive discrete numerical variables) when:

- Events happen independently
- The average rate ($\lambda$) is constant

$$X \sim Poisson(\lambda)$$

- $\lambda$ = mean = variance
 
## Posson distribution: Example $\lambda$ = 15

Suppose I get about 15 emails per workday

```{r}
#| echo: false
#| fig-align: center

x_seq <- 0:25

data.frame(x = x_seq, y = dpois(x_seq, lambda = 15)) |> 
ggplot(aes(x = x, y = y)) +
  geom_point(color = "blue") +
  geom_line(color = "blue") +
  labs(x = "Number of emails per workday")
```

Number of emails per workday $\sim Poisson(\lambda = 15)$

- Independent: Getting one email does not affect probability of getting another email
- Constant rate: average rate of emails is consistent each workday

## Posson distribution: Example $\lambda$ = 5

Now suppose I only get about 5 emails per workday

```{r}
#| echo: false
#| fig-align: center

data.frame(x = x_seq, y = dpois(x_seq, lambda = 5)) |> 
ggplot(aes(x = x, y = y)) +
  geom_point(color = "blue") +
  geom_line(color = "blue") +
  labs(x = "Number of emails per workday")
```

Number of emails per workday $\sim Poisson(\lambda = 5)$

## Changing rate $\lambda$

Poisson distribution with different $\lambda$

```{r}
#| echo: false

sim_pois <- data.frame(x = x_seq, y = dpois(x_seq, lambda = 2), lambda = 2) |> 
  bind_rows(data.frame(x = x_seq, y = dpois(x_seq, lambda = 5), lambda = 5)) |> 
  bind_rows(data.frame(x = x_seq, y = dpois(x_seq, lambda = 10), lambda = 10)) |> 
  bind_rows(data.frame(x = x_seq, y = dpois(x_seq, lambda = 15), lambda = 15)) |> 
  mutate(lambda = factor(
    lambda,
    levels = c(2, 5, 10, 15),
    labels = c("\u03bb = 2", "\u03bb = 5", "\u03bb = 10", "\u03bb = 15")
  ))
  
ggplot(sim_pois, aes(x = x, y = y)) +
  geom_point(color = "blue") +
  geom_line(color = "blue") +
  facet_grid(~lambda)
```


