---
title: "Data 200: Lecture 15a"
subtitle: "Modeling using distributions"
author: "Medina - CSU Channel Islands"
format: 
  revealjs:
    theme: [sky, ../slide-style.scss]
    slide-number: true
execute:
  echo: true 
---


```{r}
#| echo: false
options(scipen = 999)
library(janitor)
library(patchwork)
library(tidyverse)
theme_set(theme_bw(base_size = 22))
```

```{r}
#| echo: false
plot_beta_prob <- function(value, alpha, beta, direction = "lower"){
  
  data <- data.frame(x = c(0, 1))
  
  p <- ggplot(data = data, aes(x)) +
    stat_function(
      fun = stats::dbeta,
      n = 101,
      args = list(shape1 = alpha, shape2= beta)
    ) +
    labs(
      x = expression(pi),
      y = "Density"
    )
  
  if(direction == "lower"){
    
  p <- p +
      stat_function(
        fun = dbeta, 
        args = list(alpha, beta), 
        geom = "area", 
        fill = "#D55E00", 
        alpha = 0.8,
        xlim = c(0, value)
      ) +
    labs(caption = paste0(
      "The area under the curve is ", 
      round(pbeta(value, alpha, beta), digits = 4)*100, "%.", 
      " In other words, P(", "\U03C0", " < ", value, ") = ", 
      round(pbeta(value, alpha, beta), digits = 4),"."
    ))
  }
  
  
  if(direction == "upper"){
    
  p <- p +
      stat_function(
        fun = dbeta, 
        args = list(alpha, beta), 
        geom = "area", 
        fill = "#D55E00", 
        alpha = 0.8, 
        xlim = c(value, 1)
      ) +
    labs(caption = paste0(
      "The area under the curve is ", 
      round(pbeta(value, alpha, beta, lower.tail = FALSE), digits = 4)*100, "%.", 
      " In other words, P(", "\U03C0", " > ", value, ") = ", 
      round(pbeta(value, alpha, beta, lower.tail = FALSE), digits = 4),"."
    ))
  }
  
p

}


plot_normal_prob <- function(value, mean, sd, end = 3, direction = "lower"){
  
  x <- seq(mean - (end * sd) , mean + (end * sd), 0.01)
  
  y <- dnorm(x, mean = mean, sd = sd)
  
  data <- data.frame(x = x, y = y)
  
  p <- ggplot(data = data, aes(x)) +
    stat_function(
      fun = stats::dnorm,
      n = 101,
      args = list(mean = mean, sd= sd)
    ) 
  
  if(direction == "lower"){
    
    p <- p +
      stat_function(
        fun = dnorm, 
        args = list(mean, sd), 
        geom = "area", 
        fill = "#D55E00", 
        alpha = 0.8,
        xlim = c(mean - (end*sd), value)
      ) +
      labs(
        caption = paste0(
          "The area under the curve is ",
          round(pnorm(value, mean, sd), digits = 4) * 100, 
          "%."
        )
      )
  }
  
  
  if(direction == "upper"){
    
    p <- p +
      stat_function(
        fun = dnorm, 
        args = list(mean, sd), 
        geom = "area", 
        fill = "#D55E00", 
        alpha = 0.8, 
        xlim = c(value, mean+(end * sd))
      ) +
      labs(
        caption = paste0(
          "The area under the curve is ", 
          round(pnorm(value, mean, sd, lower.tail = FALSE), digits = 4) * 100, 
          "%."
        )
      )
  }
  
  p
  
}
```

## General idea

There exist known **distributions** which we know everything about.

. . .

::::{.columns}

:::{.column width="50%"}
```{r}
#| echo: false
#| fig-align: center
plot_normal_prob(value = 100, mean = 100, sd = 15) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.title.y=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()
  ) +
  labs(title = "Normal distribution")
```

::: 

:::{.column width="50%"}

:::

::::

. . .

If we have a variable in our sample that seems to follow a distribution, we can use the distribution to estimate probabilities.

## Normal distribution

```{r}
#| echo: false
#| fig-align: center

plot_normal_prob(value = 100, mean = 100, sd = 15) +
  theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.title.y=element_blank(),
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()
  )

```


This is a symmetric distribution with mean, mode, and median falling exactly in the middle. The shaded region thus shows that the area under the curve is 50% indicating that if any data follows a normal distribution, we would expect 50% of the data to fall below the mean. 


## Normal example

Consider SAT scores with a mean of 1500 and standard deviation of 300

::::{.columns}

:::{.column width="50%"}

```{r}
#| echo: false

set.seed(468)

sat_scores <- data.frame(x = rnorm(1000, mean = 1500, sd = 300))  

ggplot(sat_scores, aes(x = x)) +
  geom_histogram() +
  geom_density() +
  labs(y = "Count", x = "Sat Score", title = "Data")
```

:::

:::{.column width="50%"}

```{r}
#| echo: false


ggplot(sat_scores, aes(x = x)) +
  geom_function(fun = dnorm, args = list(mean = 1500, sd = 300), color = "blue") +
  theme(
    axis.text.y=element_blank(),
    axis.ticks.y=element_blank()
  ) +
  labs(y = "Density", x = "Sat Score", title = "Normal Model")
```

:::

::::

. . .

SAT scores are approximately normally distributed so we can approximate probabilities using the normal distribution. 

## Normal notation

The shape of the distribution is controlled by two parameters: mean ($\mu$) and standard deviation ($\sigma$).
Using notation a normal distribution is shown as $N(\mu, \sigma)$


##

```{r}
#| echo: false
#| fig-align: center
p1 <- ggplot(data = data.frame(x = c(0, 200)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 100, sd = 15)) + 
  ylab("") +
  labs(title = expression(paste("N(\u03bc = 100, ", sigma, " = ", 15, ")")))

p2 <- ggplot(data = data.frame(x = c(0, 200)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 100, sd = 30)) + 
  ylab("") +
  labs(title = expression(paste("N(\u03bc = 100, ", sigma, " = ", 30, ")")))

p3 <- ggplot(data = data.frame(x = c(0, 200)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 75, sd = 15)) + 
  ylab("") +
  labs(title = expression(paste("N(\u03bc = 75, ", sigma, " = ", 15, ")")))


p4 <- ggplot(data = data.frame(x = c(0, 200)), aes(x)) +
  stat_function(fun = dnorm, n = 1001, args = list(mean = 75, sd = 30)) + 
  ylab("") +
  labs(title = expression(paste("N(\u03bc = 75, ", sigma, " = ", 30, ")")))

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)
```

## Normal probability

```{r}
plot_normal_prob(value = 1800, mean = 1500, sd = 300, direction = "upper")
```

## Normal probability

Let's consider the SAT example again. 

- SAT scores are normally distributed with a mean of 1500 and a standard deviation of 300. 
- We can see in the previous figure that the area under the Normal curve where SAT scores are greater than 1800 is `r paste(round(pnorm(1800, 1500, 300, lower.tail = FALSE), digits = 4)*100, "%.")` 
- Thus we approximate the probability of an SAT score being greater than 1800 is `r round(pnorm(1800, 1500, 300, lower.tail = FALSE), digits = 4)`.


## Baby birth weight example

Birth weight of babies are known to be (approximately) normally distributed with a mean of 7.5 pounds and a standard deviation of 1.1 pounds. 

- If a baby is born less than 5.5 pounds then the baby is considered to be low weight. 
- What is the probability of a baby being low weight?

##

```{r}
plot_normal_prob(value = 5.5, mean = 7.5, sd = 1.1, direction = "lower")
```

The probability a baby is born low weight (<5 pounds) is approximately `r round(pnorm(5.5, 7.5, 1.1, lower.tail = TRUE), digits = 4)`.

