---
title: "Data 200: Lab 2"
subtitle: "Familiarizing yourself with data"
author: "Your name here"
format: 
  html:
    echo: true
    message: false
    warning: false
    toc: true
    embed-resources: true
---

*Please always start labs and homework by changing the author to your name. Please answer questions using complete sentences where appropriate.*


# Summarizing week 2 lecture

We will start this lab by summarizing the lecture content.
Replace each "**Fill in the blank**" with the appropriate word.


## R packages: install vs. load

-   **Fill in the blank** a package once on your computer.
-   **Fill in the blank** a package every time you start a new session / Quarto document.


## Data frames

A data frame is a rectangular table:

-   **Fill in the blank** = variables
-   **Fill in the blank** = observations (cases)

Common ways to “get to know” a dataset:

-   `glimpse()` (quick structure)
-   `head()` / `tail()` (top/bottom rows, displays 6 rows by default)
-   `nrow()` / `ncol()` (number of rows/columns)
-   `?dataset_name` or `?function_name` (to view help documentation)


## Statistical variable types

Knowing the variable type helps you choose the right summaries/plots.

Numerical

-   **Fill in the blank**: finite number of possibilities, typically counts (e.g., number of bids)
-   **Fill in the blank**: infinite number of possibilities, typically measurements (e.g., temperature)

Categorical

-   **Fill in the blank**: no natural order (e.g., US states)
-   **Fill in the blank**: natural order (e.g., rating levels, if applicable)


## Common storage types in R

Categorical storage types:

-   **Fill in the blank**, **Fill in the blank**, **Fill in the blank**

Numerical storage types:

-   **Fill in the blank**, **Fill in the blank**, **Fill in the blank**

Summarizing categorical variables

-   Counts: `count(data, var)`
-   Proportions: `tabyl(data, var)`


## Summarizing numerical variables

Measures of center:

-   **Fill in the blank**: `mean()`
-   **Fill in the blank**: `median()`

Measures of spread:

-   **Fill in the blank**: `sd()`
-   **Fill in the blank**: `var()`
-   **Fill in the blank**: `quantile()`

Use `summarize()` to compute multiple statistics at once.


# Part 1: Introduction to the data

In this lab we will be working with Wii Mario Kart Ebay auction data.
We will use functions from the tidyverse packages and data from the openintro package, so they are being loaded below.
In order to run the code chunk below both packages have to be installed on your computer.

```{r}
library(tidyverse)
library(openintro)

data("mariokart")
```


## Question 1

Use the `head()` function on the `mariokart` data in the code chunk below to view the first 6 rows of the data frame.

```{r}

```

## Question 2

Use the `glimpse()` function on the `mariokart` data to view the structure of the data.
I did not insert an R code chunk for you to use, so you will have to insert one yourself first.


## Question 3

Please look at the data documentation using `?mariokart` in the console.
For each column below please specify what type of variable it is and an appropriate R variable type.
The first one is done for you as an example.

-   `duration`: Numeric discrete
-   `id`:
-   `n_bids`:
-   `cond`:
-   `total_pr`:
-   `title`:


## Question 4

Please view the data and make sure the R storage types for each column are appropriate for the type of variable.
Note any variables that have an inappropriate R storage type.

For example the variable `start_pr` is numeric and the R storage type is "dbl", meaning `double`.
A `double` is just a numeric that can take decimals, so it is appropriate.


## Question 5

Let's practice with Github. 

- Save this file,
- Make a commit for this file with a message saying "Finished questions 1-5",
- Push your commit to GitHub
- Check GitHub online to see if you can see your changes.


# Part 2: Summarizing our data numerically

We will start part 2 on Wednesday.

I'm curious how much people ended up selling the game for.


## Question 6

Which variable contains this information?


## Question 7

Please calculate a value that tells me how much the game typically sold for.
Use inline code to answer.


## Question 8

Please calculate a value that describes the variability in final sale price.
Use inline code to answer.


## Question 9

```{r}
janitor::tabyl(mariokart, cond)
```

What percent of games were sold new?
