---
title: "Data 200: Lab 2 Solutions"
subtitle: "Familiarizing yourself with data"
author: "Your name here"
format: 
  html:
    echo: true
    message: false
    warning: false
    toc: true
    embed-resources: true
---

*Please always start labs and homework by changing the author to your name. Please answer questions using complete sentences where appropriate.*


# Summarizing week 2 lecture

We will start this lab by summarizing the lecture content.
Replace each "**Fill in the blank**" with the appropriate word.


## R packages: install vs. load

-   **Install** a package once on your computer.
-   **Load** a package every time you start a new session / Quarto document.


## Data frames

A data frame is a rectangular table:

-   **rows** = variables
-   **columns** = observations (cases)

Common ways to "get to know" a dataset:

-   `glimpse()` (quick structure)
-   `head()` / `tail()` (top/bottom rows, displays 6 rows by default)
-   `nrow()` / `ncol()` (number of rows/columns)
-   `?dataset_name` or `?function_name` (to view help documentation)


## Statistical variable types

Knowing the variable type helps you choose the right summaries/plots.

Numerical

-   **Discrete**: finite number of possibilities, typically counts (e.g., number of bids)
-   **Continuous**: infinite number of possibilities, typically measurements (e.g., temperature)

Categorical

-   **Nominal**: no natural order (e.g., US states)
-   **Ordinal**: natural order (e.g., rating levels, if applicable)


## Common storage types in R

Categorical storage types:

-   **character**, **factor**, **logical**

Numerical storage types:

-   **integer**, **double**, **numeric**

Summarizing categorical variables

-   Counts: `count(data, var)`
-   Proportions: `tabyl(data, var)`


## Summarizing numerical variables

Measures of center:

-   **mean**: `mean()`
-   **median**: `median()`

Measures of spread:

-   **standard deviation**: `sd()`
-   **variance**: `var()`
-   **quantile / percentiles**: `quantile()`

Use `summarize()` to compute multiple statistics at once.


# Part 1: Introduction to the data

In this lab we will be working with Wii Mario Kart Ebay auction data.
We will use functions from the tidyverse packages and data from the openintro package, so they are being loaded below.
In order to run the code chunk below both packages have to be installed on your computer.

```{r}
library(tidyverse)
library(openintro)

data("mariokart")
```


## Question 1

Use the `head()` function on the `mariokart` data in the code chunk below to view the first 6 rows of the data frame.

```{r}
head(mariokart)
```

## Question 2

Use the `glimpse()` function on the `mariokart` data to view the structure of the data.
I did not insert an R code chunk for you to use, so you will have to insert one yourself first.

```{r}
glimpse(mariokart)
```


## Question 3

Please look at the data documentation using `?mariokart` in the console.
For each column below please specify what type of variable it is and an appropriate R variable type.
The first one is done for you as an example.

-   `duration`: Numeric discrete
-   `id`: Nominal categorical (Just because it is stored as numbers does not mean it is numeric! These numbers don't necessarily have mathematical meaning and could just as easily been letters.)
-   `n_bids`: Numeric discrete
-   `cond`: nominal categorical
-   `total_pr`: continuous numeric (I'd also accept discrete since we know we are counting by cents at minimum)
-   `title`: nominal categorical


## Question 4

Please view the data and make sure the R storage types for each column are appropriate for the type of variable.
Note any variables that have an inappropriate R storage type.

For example the variable `start_pr` is numeric and the R storage type is "dbl", meaning `double`.
A `double` is just a numeric that can take decimals, so it is appropriate.

**Solution:**

I only noticed the `id` column as having a storage type that can be misleading since it is stored as an integer but is actually categorical.

## Question 5

Let's practice with Github. 

- Save this file,
- Make a commit for this file with a message saying "Finished questions 1-5",
- Push your commit to GitHub
- Check GitHub online to see if you can see your changes.


# Part 2: Summarizing our data numerically

We will start part 2 on Wednesday.

I'm curious how much people ended up selling the game for.


## Question 6

Which variable contains this information?

**Solution:**

`total_pr` contains the auction price plus shipping price. (Look at the documentation!)


## Question 7

Please calculate a value that tells me how much the game typically sold for.
Use inline code to answer.

**Solution:**

```{r}
mean_price <- mean(mariokart$total_pr)
```

The mean total price is `r mean_price`.


## Question 8

Please calculate a value that describes the variability in final sale price.
Use inline code to answer.

**Solution:**

```{r}
sd_price <- sd(mariokart$total_pr)
```

The standard deviation of the total price is `r sd_price`.

## Question 9

```{r}
janitor::tabyl(mariokart, cond)
```

What percent of games were sold new?

**Solution:**

41% of games were sold new.

# GitHub

When you finish this assignment don't forget to commit and push to GitHub before uploading the rendered html to Canvas.
